Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants

### 2022-03-16 paper/paper_TE_bilingual_vocabulary_model.tex
fixed typos

\linenumbers
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage[labelformat=empty]{caption}
\usepackage{caption}
\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\renewcommand{\textfraction}{.1}
\renewcommand{\floatpagefraction}{1}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}
\DeclareUnicodeCharacter{2265}{\ensuremath{\ge}}
\DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

@@

\hypertarget{accounts-of-translation-equivalent-learning}{%
\subsection{Accounts of translation equivalent learning}\label{accounts-of-translation-equivalent-learning}}

\hypertarget{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}{%
\subsubsection{Avoidance Account: Translation equivalents are harder to learn than singlets}\label{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}}

@@

\hypertarget{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}{%
\subsubsection{Preference Account: Translation equivalents are easier to learn than singlets}\label{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}}

@@

Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015). The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related). Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in children's lexicons. This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English ``dog'' and French \emph{``chien''}) and words that had other cross-language relations (e.g., ``cat'' and \emph{``chien''}). The authors then simulated bilingual vocabulaires by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals. Comparison with actual bilingual children's vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation. The authors therefore concluded that bilingual children learn translation equivalents more easily than singlets. Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat''). However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning, including vocabulary size in each language and the developmental level of a child --- a point that we will return to later in the introduction.

@@

\hypertarget{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}{%
\subsubsection{Neutral Account: Translation equivalents and singlets are learned similarly}\label{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}}

@@

\hypertarget{contributors-to-translation-equivalent-knowledge}{%
\subsection{Contributors to translation equivalent knowledge}\label{contributors-to-translation-equivalent-knowledge}}

@@

\hypertarget{vocabulary-size-in-bilinguals-two-languages}{%
\subsubsection{Vocabulary size in bilinguals' two languages}\label{vocabulary-size-in-bilinguals-two-languages}}

@@

\hypertarget{word-learnability-as-a-function-of-developmental-level}{%
\subsubsection{Word learnability as a function of developmental level}\label{word-learnability-as-a-function-of-developmental-level}}

@@

\hypertarget{the-bilingual-vocabulary-model}{%
\subsection{The Bilingual Vocabulary Model}\label{the-bilingual-vocabulary-model}}

@@

The model takes four main parameters as predictors: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS \textgreater{} 1) or against (BIAS \textless{} 1) learning translation equivalents. The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language. Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words). If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600--700 depending on the language of the adaptation. Yet, not all children will be capable of learning all of the CDI words. Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019). That is, very young children would not be expected to produce many of the advanced words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age. Since there are no CDI norms for French--English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language. For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words. Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 240 words (retrieved from the Wordbank database version 0.3.1; Frank et al., 2016). Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that averaging between the English and French CDIs at the 90th percentile provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a bilingual child of Jamie's age could potentially acquire in each language. Thus, we might expect that Jamie could potentially have learned up to 240 words in English and 240 words in French, although he has thus far only learned 50 in English and 20 in French. Note that we model learnability in terms of the number of words expected to be learnable at a particular age, rather than modeling the learnability of individual words at a particular age --- a simplifying choice that we will return to in the discussion section.

@@

\caption{\label{tab:table1}Table 1. Summary of the parameters in the Bilingual Vocabulary Model.}
\centering
\fontsize{7}{9}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}}
\toprule
Variable & Definition & Constraints & Relationship.to.other.parameters\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Main Parameters}}\\
\hspace{1em}LEARNABLE & Number of words expected to be learnable in each language, given the child’s developmental level & Varies by age. No greater than the number of words on CDI. & Maximum number that could be learned in DOM or NONDOM\\
\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (children always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM\\
\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (children always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD - DOM\\
\hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & \\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Derived Parameters}}\\
\hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & 0 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)\\
\hspace{1em}TE & Translation equivalents produced &  & TE = BIAS × DOM×NONDOM/LEARNABLE\\
\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
\hspace{1em}DOM-SINGLET & Singlets in dominant language &  & DOM-SINGLET = DOM - TE\\
\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language &  & NONDOM\_SINGLET = NONCOM - TE\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item All vocabulary measures are constrained to be integers.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

@@

\hypertarget{current-studies}{%
\subsection{Current studies}\label{current-studies}}

@@

\hypertarget{study-1-simulations}{%
\section{Study 1: Simulations}\label{study-1-simulations}}

@@

For each simulation, we generated predictions for expected patterns in bilingual children's vocabulaires. Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David \& Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016). However, others have examined derived parameters such as BALANCE (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), WORD (Legacy et al., 2016; Montanari, 2010), CONCEPT (Pearson et al., 1993), DOM-SINGLET and NONDOM-SINGLET (De Houwer et al., 2006; Pearson et al., 1993; 1995). To facilitate comparison with the extant literature where main parameters are not always reported, we generated predictions expressed both in terms of main parameters and derived parameters. This provides alternate ways of testing the model, depending on which parameters are available in the dataset. Note that predictions from the main parameters and those from derived parameters do not provide independent tests of our model for the same dataset, as they can be algebraically calculated from one another. As an analogy, a child's age is not independent from their birth date and today's date, yet we might make reference to their age in some situations and their birth date in others. Similarly, elucidating relations among both main and derived parameters provides a more complete description of expected patterns in bilingual vocabulary development.

@@

\hypertarget{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}{%
\subsection{1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}

@@

\hypertarget{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}{%
\subsection{1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}

@@

Translation equivalent knowledge was simulated across children at six developmental levels (number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM). In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation). Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM. We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE. Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE. Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model. Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will:

@@

\hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning}{%
\subsection{1.3 Simulation 3: Bias towards or against translation equivalent learning}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning}}

@@

\hypertarget{summary-of-model-predictions}{%
\subsection{1.4 Summary of model predictions}\label{summary-of-model-predictions}}

@@

\hypertarget{study-2-empirical-data}{%
\section{Study 2: Empirical data}\label{study-2-empirical-data}}

@@

\hypertarget{method}{%
\subsection{2.1 Method}\label{method}}

@@

\hypertarget{participants}{%
\subsection{2.1.1 Participants}\label{participants}}

@@

\hypertarget{measures}{%
\subsection{2.1.2 Measures}\label{measures}}

\hypertarget{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}{%
\subsubsection{MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)}\label{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}}

@@

After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1). Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).

\hypertarget{language-exposure-questionnaire-using-the-maple-approach}{%
\subsubsection{Language Exposure Questionnaire using the MAPLE approach}\label{language-exposure-questionnaire-using-the-maple-approach}}

@@

\hypertarget{procedure}{%
\subsection{2.1.3 Procedure}\label{procedure}}

@@

\hypertarget{results}{%
\subsection{2.2 Results}\label{results}}

@@

\hypertarget{descriptive-measures-of-vocabulary}{%
\subsection{2.2.1 Descriptive measures of vocabulary}\label{descriptive-measures-of-vocabulary}}

@@

Vocabulary balance (BALANCE) was then determined based on the proportion of total words produced in the non-dominant language following the formula BALANCE = NONDOM/WORD as in Study 1. On average, bilingual children in our sample had a balance score BALANCE of 0.31 (SD = 0.13), ranging from 0.02 to 0.50. SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\). The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.20\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).

Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997). We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language. For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79\%), although this was not the case for a minority of children (48 children, 21\%). The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4). The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2022, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).

@@

\hypertarget{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}{%
\subsection{2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures}\label{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

>>>>

\linenumbers
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage[labelformat=empty]{caption}
\usepackage{caption}
\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\renewcommand{\textfraction}{.1}
\renewcommand{\floatpagefraction}{1}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}
\DeclareUnicodeCharacter{2265}{\ensuremath{\ge}}
\DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
\DeclareUnicodeCharacter{2212}{\ensuremath{−}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\hypertarget{introduction}{%
\section{1 Introduction}\label{introduction}}

@@

\hypertarget{accounts-of-translation-equivalent-learning}{%
\subsection{1.1 Accounts of translation equivalent learning}\label{accounts-of-translation-equivalent-learning}}

\hypertarget{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}{%
\subsubsection{1.1.1 Avoidance Account: Translation equivalents are harder to learn than singlets}\label{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}}

@@

\hypertarget{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}{%
\subsubsection{1.1.2 Preference Account: Translation equivalents are easier to learn than singlets}\label{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}}

@@

Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015). The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related). Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in children's lexicons. This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English ``dog'' and French \emph{``chien''}) and words that had other cross-language relations (e.g., ``cat'' and \emph{``chien''}). The authors then simulated bilingual vocabularies by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals. Comparison with actual bilingual children's vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation. The authors therefore concluded that bilingual children learn translation equivalents more easily than singlets. Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat''). However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning, including vocabulary size in each language and the developmental level of a child --- a point that we will return to later in the introduction.

@@

\hypertarget{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}{%
\subsubsection{1.1.3 Neutral Account: Translation equivalents and singlets are learned similarly}\label{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}}

@@

\hypertarget{contributors-to-translation-equivalent-knowledge}{%
\subsection{1.2 Contributors to translation equivalent knowledge}\label{contributors-to-translation-equivalent-knowledge}}

@@

\hypertarget{vocabulary-size-in-bilinguals-two-languages}{%
\subsubsection{1.1.1 Vocabulary size in bilinguals' two languages}\label{vocabulary-size-in-bilinguals-two-languages}}

@@

\hypertarget{word-learnability-as-a-function-of-developmental-level}{%
\subsubsection{1.1.2 Word learnability as a function of developmental level}\label{word-learnability-as-a-function-of-developmental-level}}

@@

\hypertarget{the-bilingual-vocabulary-model}{%
\subsection{1.3 The Bilingual Vocabulary Model}\label{the-bilingual-vocabulary-model}}

@@

The model takes four main parameters as predictors: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS \textgreater{} 1) or against (BIAS \textless{} 1) learning translation equivalents. The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language. Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words). If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development Inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600--700 depending on the language of the adaptation. Yet, not all children will be capable of learning all of the CDI words. Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019). That is, very young children would not be expected to produce many of the advanced words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age. Since there are no CDI norms for French--English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language. For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words. Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 240 words (retrieved from the Wordbank database version 0.3.1; Frank et al., 2016). Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that averaging between the English and French CDIs at the 90th percentile provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a bilingual child of Jamie's age could potentially acquire in each language. Thus, we might expect that Jamie could potentially have learned up to 240 words in English and 240 words in French, although he has thus far only learned 50 in English and 20 in French. Note that we model learnability in terms of the number of words expected to be learnable at a particular age, rather than modeling the learnability of individual words at a particular age --- a simplifying choice that we will return to in the discussion section.

@@

\caption{\label{tab:table1}Table 1. Summary of the parameters in the Bilingual Vocabulary Model.}
\centering
\fontsize{7}{9}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}}
\toprule
Variable & Definition & Constraints & Relationship.to.other.parameters\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Main Parameters}}\\
\hspace{1em}LEARNABLE & Number of words expected to be learnable in each language, given the child’s developmental level & Varies by age. No greater than the number of words on CDI. & Maximum number that could be learned in DOM or NONDOM\\
\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (children always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1 – BALANCE)×WORD; DOM = WORD – NONDOM\\
\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (children always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD – DOM\\
\hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & \\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Derived Parameters}}\\
\hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & 0 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1 – BALANCE); WORD = NONDOM/(BALANCE)\\
\hspace{1em}TE & Translation equivalents produced &  & TE = BIAS × DOM×NONDOM/LEARNABLE\\
\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD – TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
\hspace{1em}DOM-SINGLET & Singlets in dominant language &  & DOM-SINGLET = DOM – TE\\
\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language &  & NONDOM-SINGLET = NONCOM – TE\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item All vocabulary measures are constrained to be integers.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

@@

\hypertarget{current-studies}{%
\subsection{1.4 Current studies}\label{current-studies}}

@@

\hypertarget{study-1-simulations}{%
\section{2 Study 1: Simulations}\label{study-1-simulations}}

@@

For each simulation, we generated predictions for expected patterns in bilingual children's vocabularies. Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David \& Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016). However, others have examined derived parameters such as BALANCE (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), WORD (Legacy et al., 2016; Montanari, 2010), CONCEPT (Pearson et al., 1993), DOM-SINGLET and NONDOM-SINGLET (De Houwer et al., 2006; Pearson et al., 1993; 1995). To facilitate comparison with the extant literature where main parameters are not always reported, we generated predictions expressed both in terms of main parameters and derived parameters. This provides alternate ways of testing the model, depending on which parameters are available in the dataset. Note that predictions from the main parameters and those from derived parameters do not provide independent tests of our model for the same dataset, as they can be algebraically calculated from one another. As an analogy, a child's age is not independent from their birth date and today's date, yet we might make reference to their age in some situations and their birth date in others. Similarly, elucidating relations among both main and derived parameters provides a more complete description of expected patterns in bilingual vocabulary development.

@@

\hypertarget{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}{%
\subsection{2.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}

@@

\hypertarget{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}{%
\subsection{2.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}

@@

Translation equivalent knowledge was simulated across children at six developmental levels (number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM). In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation). Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM. We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM -- TE and NONDOM-SINGLET = NONDOM -- TE. Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE. Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model. Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will:

@@

\hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning}{%
\subsection{2.3 Simulation 3: Bias towards or against translation equivalent learning}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning}}

@@

\hypertarget{summary-of-model-predictions}{%
\subsection{2.4 Summary of model predictions}\label{summary-of-model-predictions}}

@@

\hypertarget{study-2-empirical-data}{%
\section{3 Study 2: Empirical data}\label{study-2-empirical-data}}

@@

\hypertarget{method}{%
\subsection{3.1 Method}\label{method}}

@@

\hypertarget{participants}{%
\subsubsection{3.1.1 Participants}\label{participants}}

@@

\hypertarget{measures}{%
\subsection{3.1.2 Measures}\label{measures}}

\hypertarget{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}{%
\subsubsection{3.1.2.1 MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)}\label{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}}

@@

After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM -- TE and NONDOM -- TE as in Study 1). Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD -- TE as in Study 1).

\hypertarget{language-exposure-questionnaire-using-the-maple-approach}{%
\subsubsection{3.1.2.2 Language Exposure Questionnaire using the MAPLE approach}\label{language-exposure-questionnaire-using-the-maple-approach}}

@@

\hypertarget{procedure}{%
\subsection{3.1.3 Procedure}\label{procedure}}

@@

\hypertarget{results}{%
\subsection{3.2 Results}\label{results}}

@@

\hypertarget{descriptive-measures-of-vocabulary}{%
\subsection{3.2.1 Descriptive measures of vocabulary}\label{descriptive-measures-of-vocabulary}}

@@

Vocabulary balance (BALANCE) was then determined based on the proportion of total words produced in the non-dominant language following the formula BALANCE = NONDOM/WORD as in Study 1. On average, bilingual children in our sample had a balance score BALANCE of 0.31 (SD = 0.13), ranging from 0.02 to 0.50. Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\). The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.20\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).

Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Hurtado et al., 2014; Pearson et al., 1997). We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language. For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79\%), although this was not the case for a minority of children (48 children, 21\%). The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4). The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).

@@

\hypertarget{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}{%
\subsection{3.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures}\label{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}}

@@

\hypertarget{discussion}{%
\section{4 Discussion}\label{discussion}}

@@

\hypertarget{conclusions}{%
\section{5 Conclusions}\label{conclusions}}
